\section{Discussion}

% Discussion: Discuss the assignment and your achievements.
% You are free to critically assess your work – what could have been done better, which way you would choose to go if given the same task again etc.
% You can also provide feedback about the assignment itself.

% Here comes discussion. How did we do?

%TODO:  ville ha en liten introduksjon så vi ikke starter et kapittel med å starte et underkapittel
%       4. Discussion
%       4.1 Req
%       synes det er bedre om det er noe i mellom der

This section discusses the test results and performance of the presented solution,
as well as the work process behind implementing it.

\subsection{Requirements}
All the requirements given in the exercise were met.
The processor implements a basic subset of the MIPS instruction set \cite[p.64]{compendium} as specified.
It has been tested on a top-level and each component is also tested in separation.
The processor is also successfully tested on the FPGA using the hostcomm \cite{hostcomm} utility.

\subsection{Performance}
% TODO: det er en del ting som sies på nytt her, som ikke er så veldig bra, burde skrives om
Because of all the precausions done with performance in mind, the final product has a very high performance.
The maximum achieved clock speed is 73.573MH for the processor.
This is significantly better than the limit of 24 MHz set by the testing framework for the FPGA.

To make sure the performance is as good as possible, a lot of choices have been made during the development.
Most importantly, the fact that Xilinx does a lot of work in the optimizing department has been taken into concideration.
This means that the code is largely written to make sure Xilinx understands what the code is describing.
One example of this is the elimination of self written state values.
These are instead made as an enum which in terms make Xilinx change the values to whatever is best for the situation.

\subsection{Energy Efficiency}
There aren't many ways to make the design very energy efficient.
The communication with the external memory only takes one cycle, so the processor always has something to do.
If, however, the communication would take multiple cycles, the rest of the processor could sleep while waiting for the memory.

Another way to make the design more energy efficient is to make sure all parts of the processor are used all the time.
If every part of the processor are being used every clock cycle,
the total amount of clock cycles is reduced and less power is consumed.
A common implementation of this is known as a pipelining processor, and will be looked at in the next exercise.

\subsection{The work process}

From the start, we focused on creating a minimal working processor which fulfilled all the basic requirements given.
As the processor was naturally divided into multiple quite independent components with well-defined interfaces, we managed to parallelize the development to some extent.
This gave us a fast development speed in the beginning.

The biggest challenge was when all components were deemed ready (passing their testbenches), and were wired together in the toplevel processor.
It turned out not everything worked as expected, and the toplevel design would have to be carefully inspected.
A way that this could have gone more smoothly was if there were more higher-level testbenches implemented by us, for example testing 3 or 4 components together.

When the toplevel testbench finally passed, we tackled the task of flashing the program to the FPGA.
Except for some minor issues with project setup in Xilinx with regards to the DualPortMem module, hostcomm made this a quite simple task.

\subsection{Further work}

There is still room for optimization, to further increase the clock speed.
Another interesting improvement would be to implement support for a wider range of the MIPS instruction set.
Having support for all the defined instructions, would for instance enable us to run all sorts of available MIPS programs on our processor.
